---
title: "Assignment 6"
author: "edelsonc"
date: "September 11, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("/Users/edelsonc/Desktop/Data_Science/EDA/assignment_6")
```
This assignment will investigate descritizing data in R. This will be done using a data set looking as Chronic Kidney Disease, provided from the University of California Irvine.

First the data is downloaded and extracted using bash.
```{r, engine= "bash", eval=FALSE}
# file url
url="https://archive.ics.uci.edu/ml/machine-learning-databases/00336/Chronic_Kidney_Disease.rar"

# use wget command to download the file
wget "$url"

# extract the file's content
unrar e Chronic_Kidney_Disease.rar

# record the data providence
echo "Downloaded from $url at $( date )" > KidneyProvidence.txt
```

Now lets investigate the extracted files

```{r, engine="bash"}
# list content of current directory, looking for items with chronic in
# their name
ls | grep chronic
```

The data files have been extracted. However, they are in an unusual format, a `.arff` document. In order to read this exotic file into memory, we'll need to use a special package, `RWeka`

```{r}
# load RWeka library
library(RWeka)

# use arff reading function to load data into memory
kidney <- read.arff("chronic_kidney_disease_full.arff")
str(kidney)
```


## Completeness of data

We can investigate how many rows are complete using the `complete.cases` command

```{r}
sum(complete.cases(kidney))/nrow(kidney) * 100
```

So `r sum(complete.cases(kidney))/nrow(kidney) * 100`% or the rows contain an entry for every column.

We wish to investigate the potassium column, and inorder to do this we'll eliminate rows where there is an `NA` for potassium

```{r}
kidney<- kidney[!(is.na(kidney$pot)),]
```

## Descritizing the Data

There are a number of different schreme for descritizing data. A common one is to simply break the data into n many chunks, each with the same number of points in it

```{r}
bin_loc <- quantile(kidney$pot, probs= seq(0,1, by=0.2))

kidney$dis_fix_freq <- with(kidney, cut(pot, breaks=bin_loc, include.lowest=TRUE, right=FALSE))

summary(kidney$dis_fix_freq)
```

An alternate scheme for descretizing is to divide the range into equal parts

```{r}
MIN <- range(kidney$pot)[1]
MAX <- range(kidney$pot)[2]

bin_loc <- seq(MIN, MAX, (MAX - MIN)/5)

kidney$dis_width <- with(kidney, cut(pot, breaks=bin_loc, include.lowest=TRUE, right=FALSE))

summary(kidney$dis_width)
```

We can see that for this example, this is not the best discretization.

An alternate regime is to use a topdown discretization process. The `R` package `discretization` lets you do that using the function `disc.Topdown` with anyone of three algorithims: CAIM, CACC, Ameva.

```{r, warning=FALSE}
library("discretization")

kidney_comp <- kidney[complete.cases(kidney),]

top_down <- disc.Topdown(kidney_comp[c(13,14)], method=1)

kidney_comp$top_down <- factor(top_down$Disc.data[[c(2)]])

summary(kidney_comp$top_down)
```

Alternatively, a bottom up approach can be used
```{r, warning=FALSE}
bottom_up <- chiM(kidney_comp[c(13, 14)])

kidney_comp$bottom_up <- factor(bottom_up$Disc.data[[c(2)]])

summary(kidney_comp$bottom_up)