---
title: "Assignment_1"
author: "edelsonc"
date: "August 23, 2016"
output: pdf_document
---

# Introduction
This R-markdown document will serve as assignment 1. In it I will first make use of the R-markdown language to write a PDF. Then I will import a csv and a json into RStudio.

# Data Aquisition
All data used was aquired from data.gov. Specifically, the data set *Somerville Police Officer Crash Report 2010-August 2015* was used.

### Reading the CSV
First, the data was read as a csv file from the following
```{r eval=FALSE}
csv_url <- "https://data.somervillema.gov/api/views/jdcq-3gza/rows.csv?accessType=DOWNLOAD"
```
This is then easily read into memory with the download command

```{r eval=FALSE}
download.file(csv_url, "~/MyDir/Crash.csv")
```
where ```MyDir``` can be anypath of choice.

### Reading from JSON
Reading from a JSON isn't particulary more difficult. However, it does require a new R library. There are three to choose from: ```rjson```, ```rjsonlite```, or ```RJSONIO```. For our purposes, we will use ```jsonlite```.

First the rjsonlite library must be loaded
```{r eval=FALSE}
library("jsonlite")
```

Then we specify the url to download the JSON from
```{r eval=FALSE}
json_url <- "https://data.somervillema.gov/api/views/jdcq-3gza/rows.json?accessType=DOWNLOAD"
```

At this point we use the 'fromJSON' to conviniently read the file into RStudio
```{r eval=FALSE}
JSONData <- fromJSON(json_url)
```