---
title: "Assignment 2"
author: "edelsonc"
date: "August 28, 2016"
output: pdf_document
---

This exercise will deal with extensible markdown files (xmls). Xml files are very similar to html files in that they have a tag structure that encasulates information. However, as opposed to html, xml allows users to create their own tags, hence deriving its name of extensible. Because of this, a lot of information can be gleaned from xml files.

## Downloading xmls
For this exercise, an xml file from Reed college was used. This was downloaded using the now familar download command:
``` {r eval=FALSE}
# the url will be saved and used to both download the xml and recorded the data
# providence
xml_url <- "http://www.cs.washington.edu/research/xmldatasets/data/courses/reed.xml"
xml_file <- "~/Desktop/Data_Science/EDA/assignment_2/reed.xml"
download.file(xml_url, xml_file)

# Data providence is recorded and saved in a separate text file for later reference
ReedCoursesProvidence <- <- paste("Downloaded from", xml_url, "At", Sys.time(), sep = " ")
write(ReedCoursesProvidence, "~/Desktop/Data_Science/EDA/assignment_2/ReedProvidence.txt")
```

Now the data is ready to be parsed.

## Parsing the xml
Since xml is such a prevelant file type, unsurprisingly there exists a library for dealing with it named XML. XML comes with a number of useful functions, such as the ```xmlToList ```, which we can use to convert our xml file into an R list.
``` {r eval=FALSE}
# load the XML library to be used to convert out new xml file to a list
library("XML)
xml_data <- xmlToList(xml_file)
```

However, this has a problem. This list contains a number of entries which are ```NULL```. ```NULL``` values are automatically dropped from vectors or matrices, making converting our new list into a matrix or dataframe difficult. To avoid this, we'll instead use the ```xmlToDataFrame``` function, which handles these tricky cases for us
```{r eval=FALSE}
xml_dataframe <- xmlToDataFrame(xml_file)
```

## Counting Values
Counting fields is a common problem in most programming languages. Unsuprisingly, ```R``` has a number of useful functions to count fields. The first of these is ```length```. It returns the length of a given vector or list.

``` {r eval=FALSE}
length(xml_data)  # returns 703
length(xml_dataframe$title) # also returns 703
```

If we want to know how many unique instances of a field there are, there is a simple function to allow that, ```unique```. ```unique``` returns a list or vector where it has removed all replicate enteries. Thus we can use ```length``` and ```unique``` together in order to learn how many distinct courses are offered at Reed College
``` {r eval=FALSE}
# using the unique command with the length first creates a list of titles only
# containing unique values and then returns the length of the list
num_courses <- length(unique(xml_dataframe$title))  # returns 394
```

Counting the number of courses with a ```NULL``` instructor is a little more difficult. When ```xmlToDataFrame``` was applied to our reed.xml file, it took empty fields (ie ```""```) and directly stuck them into the dataframe as is. Therefore, to learn how many courses has no instructor listed, we must count the number of instructors listed as ```""```.

A simplistic way of accomplishing this is to use the ```sapply``` function. This function takes a vector or a list and applied a function to every element of the list. Using it, we can check if an element is ```""``` and then use the builtin ```sum``` to count how many times it occured

``` {r eval=FALSE}
# use the sapply function to return a boolean vector
isnull <- sapply(xml_dataframe$instructor, is.element, el="")

# applying sum to a boolean returns the number of TRUEs since TRUE=1
empty_instructor <- sum(isnull)  # returns 15
```

Finally, the number of distinct instuctors can be aquired in the same way as distinct courses, making sure to subtract 1 for the empty field
```{r eval=FALSE}
num_instructors <- length(unique(xml_dataframe$instructor)) - 1  # 135
```